---
title: "Supplementary materials | Investigating Glottal Stop Coda Loss During Sound Change of Checked Syllables Based on Speech-EGG Voice Offset Alignment"
author: "Bingliang Zhao"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      options(warn=-1))
library(readxl)
library(writexl)
library(dplyr)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(ggExtra)
library(ggpubr)
library(ggpmisc)
library(patchwork)
library(factoextra)
library(forcats)
library(FactoMineR)
library(lme4)
library(lmerTest)
library(export)
library(showtext)
library(Hmisc)
library(broom)
library(ggridges)
library(foreign)
library(agricolae)

# font setting
font_add("arial","arial.ttf")
showtext_auto()
```

# Get ready

Data from four female participants (OF8, OF9, YF6, YF7) and one male participant (OM6) were excluded due to the poor quality of their EGG signals.

Audio files with incidental noise were also excluded.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
# load the original database
load('ZhaoInterspeech2025.RData')

# Marking the outliers for every age*sex group using the 3-standard deviation criterion
df.ALL = df.interspeech2025data %>%
  group_by(age,sex) %>%
  mutate(VOL_mean = mean(VOL,na.rm = TRUE),
         VOL_sd = sd(VOL,na.rm = TRUE),) %>%
  ungroup()
df.ALL$outlier = ifelse(df.ALL$VOL >= df.ALL$VOL_mean-3*df.ALL$VOL_sd & df.ALL$VOL <= df.ALL$VOL_mean+3*df.ALL$VOL_sd,0,1)

# 3,027 valid tokens for subsequent analysis
df.VALID = filter(df.ALL,outlier == 0)
```

Calculate R^2^ of CQ & VOL for every participant.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
participants = unique(df.ALL$participant)
paras = c('CQ', 'VOL')
df.R2 = data.frame(participants)
colnames(df.R2)[1] = 'participant'
colnames(df.participantInfo)[1] = 'participant'
df.R2$age = ifelse(substr(as.character(df.R2$participant),1,1) == "O", 'Old','Young')
df.R2$sex = ifelse(substr(as.character(df.R2$participant),2,2) == "M", 'Male','Female')
df.R2 = merge(df.R2,df.participantInfo,by = 'participant')
```

Run GLMs for every participant
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
for (i in c(1:length(participants))){
  for (para in paras){
    m = lm(get(para) ~ syllable_type,data = filter(df.ALL,participant == participants[i]))
    # dependent variable: CQ / VOL
    # fixed factor: syllable type (binary coded)
    summary_model = summary(m)
    p_value = summary_model$coefficients[2, 4]
    beta = summary_model$coefficients[2, 1]
    r2_adj = summary_model$adj.r.squared
    df.R2[i,paste(para,'p',sep='_')] = p_value
    df.R2[i,paste(para,'b',sep='_')] = beta
    df.R2[i,paste(para,'r',sep='_')] = r2_adj
    if (para == 'VOL'){
      df.R2[i,paste(para,'Yes',sep='_')] = ifelse(p_value < 0.05 & beta<0,1,0)
    }
    else{
      df.R2[i,paste(para,'Yes',sep='_')] = ifelse(p_value < 0.05 & beta>0,1,0)
    }
  }
}
```

As checked syllables are expected to have more glottal constriction, higher CQ is anticipated for checked. Therefore, CQ_Yes is a binary factor whose value is equal to 1 when checked CQ is significantly higher than unchecked CQ (beta > 0 & p < .05).

VOL_Yes is a binary factor whose value is equal to 1 when checked VOL is significantly lower than unchecked CQ (beta < 0 & p < .05).
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
df.R2$CQ_Yes = factor(df.R2$CQ_Yes)
df.R2$VOL_Yes = factor(df.R2$VOL_Yes)
```

# Statistical analysis

## Section 3.1

The R^2^s, beta coefficients, and _p_ values of CQ/VOL generated from GLMs for every participant can be seen in **df.R2** (visualized in the following figure).
```{r}
knitr::kable(df.R2)
```

```{r echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}
p.PPT = ggplot(df.R2, aes(x = CQ_r, y = VOL_r, color = interaction(CQ_Yes, VOL_Yes)))+
  geom_point(size=3)+
  scale_color_manual(values = c('black','#2c7bb6','#c0321a','#fdae61')) +
  theme_bw(base_family = "arial")+
  labs(x = expression(paste(R^2,' for CQ')),y = expression(paste(R^2,' for VOL')))+
  theme(strip.text = element_text(size = 16,hjust=0),
        strip.background = element_blank(),
        plot.title = element_text(size = 18, face = "bold"),
        axis.title.x = element_text(size = 18, face = "bold"),
        axis.title.y = element_text(size = 18, face = "bold", angle = 90),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 16),
        legend.position = "bottom",
        panel.border = element_rect(colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "transparent"))+facet_wrap(~age)

p.PPT 
```

A strong positive correlation was discovered between R^2^ values of VOL and CQ. All participants were included in the Pearson's r test.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
cor.test(df.R2$CQ_r,df.R2$VOL_r)
```

## Section 3.2

### Age-related differences in raw VOL

Linear mixed-effects models were employed to examine VOL differences between checked and unchecked syllables. Age & sex fixed effects and their interactions with syllable type were tested.

We first fitted the base model which included fixed effects of syllable_type/age/sex, their two-way/three-way interactions and random intercept of participant.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
m.base = lmer(VOL~syllable_type*age*sex + (1|participant), df.VALID)
```

Including random intercept of item significantly improved model fit compared to **m.base** (p < .05).
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
m.itemAdd = lmer(VOL~syllable_type*age*sex + (1|participant) + (1|item), df.VALID)
anova(m.base,m.itemAdd)
```

Including random slope of syllable_type for participant significantly improved model fit compared to **m.itemAdd** (p < .05).
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
m.typeAnditemAdd = lmer(VOL~syllable_type*age*sex + (syllable_type|participant) + (1|item), df.VALID)
anova(m.itemAdd, m.typeAnditemAdd)
```

Including random slope of age/sex for item didn't improve model fit compared to **m.itemAdd** (p > .05).
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
m.itemAndageAdd = lmer(VOL~syllable_type*age*sex + (1|participant) + (age|item), df.VALID)
m.itemAndsexAdd = lmer(VOL~syllable_type*age*sex + (1|participant) + (sex|item), df.VALID)
anova(m.itemAdd, m.itemAndageAdd, m.itemAndsexAdd)
```

Including random slope of age&sex for item led to boundary (singular) fit or failure to converge.
```{r echo = TRUE, warning = TRUE, error = TRUE, message = TRUE}
m.itemAndageAndsexAdd = lmer(VOL~syllable_type*age*sex + (1|participant) + (age+sex|item), df.VALID)
m.itemAndagePlussexAdd = lmer(VOL~syllable_type*age*sex + (1|participant) + (age*sex|item), df.VALID)
```

compared to **m.typeAnditemAdd**, including random slope of age for item significantly improved model fit (p < .05) while including random slope of sex for item didn't.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
m.itemAndageWithtypeAdd = lmer(VOL~syllable_type*age*sex + (syllable_type|participant) + (age|item), df.VALID)
m.itemAndsexWithtypeAdd = lmer(VOL~syllable_type*age*sex + (syllable_type|participant) + (sex|item), df.VALID)
anova(m.typeAnditemAdd, m.itemAndageWithtypeAdd, m.itemAndsexWithtypeAdd)
```

Again,including random slope of age&sex for item led to boundary (singular) fit.
```{r echo = TRUE, warning = TRUE, error = TRUE, message = TRUE}
m.itemAndsexAndageWithtypeAdd = lmer(VOL~syllable_type*age*sex + (syllable_type|participant) + (age+sex|item), df.VALID)
```

Finally, **m.itemAndageWithtypeAdd** was chosen as the best model. Here are the outputs.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
summary(m.itemAndageWithtypeAdd)
```

### Age-related differences in mean and S.D. of VOL

The mean and standard deviation of VOL for checked and unchecked syllables were separately calculated for each participant.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
df.SUM = df.VALID %>%
  group_by(participant,syllable_type,age,sex) %>%
  summarise(VOL_mean = mean(VOL,na.rm = TRUE),
            VOL_sd = sd(VOL,na.rm = TRUE)) %>%
  ungroup()
```

In the older group, both the mean and standard deviation of VOL for checked syllables were significantly smaller than those for unchecked syllables.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
m.VOLmean.Old = lm(VOL_mean ~ syllable_type, filter(df.SUM,age == 'Old'))
summary(m.VOLmean.Old)
m.VOLsd.Old = lm(VOL_sd ~ syllable_type, filter(df.SUM,age == 'Old'))
summary(m.VOLsd.Old)
```

In the younger group, these measures showed no significant difference between checked and unchecked syllables.
```{r echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
m.VOLmean.Young = lm(VOL_mean ~ syllable_type, filter(df.SUM,age == 'Young'))
summary(m.VOLmean.Young)
m.VOLsd.Young = lm(VOL_sd ~ syllable_type, filter(df.SUM,age == 'Young'))
summary(m.VOLsd.Young)
```

# Graphs

## Figure 2

Figure2_left shows the correlation of VOL with CQ among the eighteen participants whose checked CQ is significantly higher than unchecked CQ.

Figure2_right shows the distributions of VOL/CQ for an old male participant (OM8)
```{r echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}
Figure2_left = ggplot(filter(df.R2,CQ_Yes == 1), aes(x = CQ_r, y = VOL_r))+
  geom_smooth(method = lm,color = 'gray80',fill = 'gray80',alpha = 0.2,linewidth = 1)+
  geom_point(aes(shape = sex,color = age),size = 2.5)+
  annotate("segment", x = 0.756, y = 0.34, xend = 0.756, yend = 0.375, arrow = arrow(length=unit(0.2,"cm")),linetype=1,size=0.8, color = "#c0321a")+
  scale_color_manual(values = c('black','gray60')) +
  scale_shape_manual(values = c(16,17))+
  labs(x = expression(paste(R^2,' for CQ')),y = expression(paste(R^2,' for VOL')))+
  theme_bw(base_family = "arial")+
  theme(strip.text = element_text(size = 16,hjust=0),
        strip.background = element_blank(),
        plot.title = element_text(size = 18, face = "bold"),
        axis.title.x = element_text(size = 18, face = "bold"),
        axis.title.y = element_text(size = 18, face = "bold", angle = 90),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 16),
        legend.position = "bottom",
        panel.border = element_rect(colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "transparent"))+guides(shape = guide_legend(ncol = 1),color = guide_legend(ncol = 1))
Figure2_rightCQ = ggplot(filter(df.VALID, participant == 'OM8'), aes(x= CQ,fill = syllable_type))+
  geom_boxplot(width=0.04,position=position_dodge(width=0.045),staplewidth = 0.5)+
  scale_fill_manual(name = "Syllable types", 
                    labels = c("Unchecked", "Checked"), 
                    values = c('#2c7bb6','#fdae61')) +
  labs(x='CQ (%)')+
  theme_bw(base_family = "arial")+
  theme(strip.text = element_text(size = 16,hjust=0),
        strip.background = element_blank(),
        plot.title = element_text(size = 18, face = "bold"),
        axis.title.x = element_text(size = 18, face = "bold"),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 16),
        legend.position = "bottom",
        panel.border = element_rect(colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "transparent"))+guides(fill = guide_legend(ncol = 1))
Figure2_rightVOL = ggplot(filter(df.VALID, participant == 'OM8'), aes(x= VOL,fill = syllable_type))+
  geom_boxplot(width=0.04,position=position_dodge(width=0.045),staplewidth = 0.5)+
  scale_fill_manual(name = "Syllable types", 
                    labels = c("Unchecked", "Checked"), 
                    values = c('#2c7bb6','#fdae61')) +
  labs(x='VOL (ms)')+
  theme_bw(base_family = "arial")+
  theme(strip.text = element_text(size = 16,hjust=0),
        strip.background = element_blank(),
        plot.title = element_text(size = 18, face = "bold"),
        axis.title.x = element_text(size = 18, face = "bold"),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 16),
        legend.position = "bottom",
        panel.border = element_rect(colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "transparent"))+guides(fill = guide_legend(ncol = 1))
Figure2_right = (Figure2_rightVOL + Figure2_rightCQ +plot_layout(nrow=2)) + plot_layout(guides = 'collect')&theme(legend.position = 'bottom')
Figure2 = Figure2_left + Figure2_right
Figure2
```

## Figure 3

```{r echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}
Figure3 = ggplot(df.VALID, aes(x= VOL, fill = syllable_type))+
  geom_density(alpha = 0.5,linewidth = 0.5,adjust = 1,position = 'identity')+
  geom_boxplot(aes(y = -0.03),width=0.04,position=position_dodge(width=0.045),staplewidth = 0.5)+
  geom_vline(aes(xintercept=0),linetype="dashed")+
  scale_fill_manual(name = "Syllable types", 
                    labels = c("Unchecked", "Checked"), 
                    values = c('#2c7bb6','#fdae61')) +
  labs(x='VOL (ms)')+
  theme_bw(base_family = "arial")+
  theme(strip.text = element_text(size = 16,hjust=0),
        strip.background = element_blank(),
        plot.title = element_text(size = 18, face = "bold"),
        axis.title.x = element_text(size = 18, face = "bold"),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 16),
        legend.position = "bottom",
        panel.border = element_rect(colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "transparent")) + facet_grid(sex~age)
Figure3
```

## Figure 4
```{r echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}
Figure4_left = ggplot(df.SUM, aes(y= VOL_mean,x = age,fill = syllable_type))+
  geom_boxplot(staplewidth = 0.5)+ 
  labs(y = 'Mean of VOL (ms)')+
  scale_fill_manual(name = "Syllable types", 
                    labels = c("Unchecked", "Checked"), 
                    values = c('#2c7bb6','#fdae61')) +
  theme_bw(base_family = "arial")+
  theme(strip.text = element_text(size = 16,hjust=0),
        strip.background = element_blank(),
        plot.title = element_text(size = 18, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 18, face = "bold", angle = 90),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 16),
        legend.position = "bottom",
        panel.border = element_rect(colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "transparent"))

Figure4_right = ggplot(df.SUM, aes(y= VOL_sd,x = age,fill = syllable_type))+
  geom_boxplot(staplewidth = 0.5)+
  labs(y='S.D. of VOL (ms)')+
  scale_fill_manual(name = "Syllable types", 
                    labels = c("Unchecked", "Checked"), 
                    values = c('#2c7bb6','#fdae61')) +
  theme_bw(base_family = "arial")+
  theme(strip.text = element_text(size = 16,hjust=0),
        strip.background = element_blank(),
        plot.title = element_text(size = 18, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 18, face = "bold", angle = 90),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 16),
        legend.position = "bottom",
        panel.border = element_rect(colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "transparent"))

Figure4 = Figure4_left + Figure4_right +
  plot_layout(guides = 'collect')&theme(legend.position = 'bottom')

Figure4
```
